---
title: "MiniProject1"
author: "Ojaas Hampiholi"
date: "2/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidycensus)
library(rio)
library(broom)
library(mgcv)
library(dplyr)
library(gridExtra)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73",
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

Key is dcede92dc65722d616da79d0a47a759893f274e4

Population1990 %>%
  rename(value = Population1990)

```{r}
#census_api_key("dcede92dc65722d616da79d0a47a759893f274e4")
Population1990 <- get_decennial(geography = "state", variables = "P0010001", year = 1990)
Population1990 = as.data.frame((Population1990))
Population1990 <- Population1990[c(2,4)]
names(Population1990)[names(Population1990) == "value"] <- "Population1990"
head(Population1990)
#View(Population1990)
```

```{r}
Population2000 <- get_decennial(geography = "state", variables = "P001001", year = 2000)
Population2000 = as.data.frame(Population2000)
names(Population2000)[names(Population2000) == "value"] <- "Population2000"
Population2000 = as.data.frame((Population2000))
Population2000 <- Population2000[c(2,4)]

head(Population2000)
#View(Population2000)
```

```{r}
Population2010 <- get_decennial(geography = "state", variables = "P001001", year = 2010)
Population2010 = as.data.frame(Population2010)
names(Population2010)[names(Population2010) == "value"] <- "Population2010"
Population2010 = as.data.frame((Population2010))
Population2010 <- Population2010[c(2,4)]
head(Population2010)
#View(Population2010)

Population2018 <- get_estimates(geography = "state", product = "population")
Population2018 = as.data.frame(population2018)
names(Population2018)[names(Population2018) == "value"] <- "Population2018"
Population2018 <- Population2018[c(1,4)]
Population2018 = Population2018[1:52,]
head(Population2018)
#View(Population2018)
```

```{r}
state_abbrev = as.data.frame(read.delim(file = "state_abbrevs.txt", header = TRUE, sep = " "))
state_abbrev = data.frame(lapply(state_abbrev, as.character), stringsAsFactors=FALSE)
names(state_abbrev)[names(state_abbrev)=="State"] <- "NAME"
head(state_abbrev)
#View(state_abbrev)
```

```{r}
populationChanges = inner_join(state_abbrev, Population1990, by = "NAME")
populationChanges = inner_join(populationChanges, Population2000, by = "NAME")
populationChanges = inner_join(populationChanges, Population2010, by = "NAME")
populationChanges = inner_join(populationChanges, Population2018, by = "NAME")
#View(populationChanges)


head(populationChanges)
#View(populationChanges)
```


```{r, fig.height=8}

#populationChanges$change1018 = populationChanges$Population2018 - populationChanges$Population2010
#populationChanges$change0010 = populationChanges$Population2010 - populationChanges$Population2000
#populationChanges$change9000 = populationChanges$Population2000 - populationChanges$Population1990

#populationChanges$changePercent1018 = (populationChanges$Population2018 - populationChanges$Population2010) / #populationChanges$Population2010
#populationChanges$changePercent0010 = (populationChanges$Population2010 - populationChanges$Population2000) / #populationChanges$Population2000
#populationChanges$changePercent9000 = (populationChanges$Population2000 - populationChanges$Population1990) / #populationChanges$Population1990


ggplot(populationChanges, aes(x = reorder(Code, change0010), y = value, color = ColorsIndicate)) +
    geom_point(aes(y = change9000, col = "change9000")) + 
    geom_point(aes(y = change0010, col = "change0010")) +
    geom_point(aes(y = change1018, col = "change1018")) +
  facet_wrap(Region ~ .) +
  #geom_smooth(method="gam", formula = y~s(x), se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Population Density Change by Region") +
  xlab("State") +
  ylab("Population Change")
```

```{r, fig.height=8}
ggplot(populationChanges, aes(x = reorder(Code, changePercent1018), y = value, color = ColorsIndicate)) +
    geom_point(aes(y = changePercent9000, col = "changePercent9000")) + 
    geom_point(aes(y = changePercent0010, col = "changePercent0010")) +
    geom_point(aes(y = changePercent1018, col = "changePercent1018")) +
  facet_wrap(Region ~ .) +
  #geom_smooth(method="gam", formula = y~s(x), se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Percent Population Density Change by Region") +
  xlab("State") +
  ylab("Population Change")
```

```{r}
ggplot(populationChanges, aes(x =  reorder(Code, change9000), y = value, color = ColorsIndicate)) + 
    geom_point(aes(y = change9000, col = "change9000")) + 
    geom_point(aes(y = change0010, col = "change0010")) +
    geom_point(aes(y = change1018, col = "change1018")) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  
```




```{r}
jan1990 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 1990, ]
dec1999 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 1999, ]
dec1999$Difference = dec1999$Price - jan1990$Price


jan2000 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 2000, ]
dec2009 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 2009, ]
dec2009$Difference = dec2009$Price - jan2000$Price

jan2010 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 2010, ]
dec2018 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 2018, ]
dec2018$Difference = dec2018$Price - jan2010$Price

```
  
  
  
```{r}
newHousingData = inner_join(dec1999, dec2009, by = "Code")
newHousingData = inner_join(newHousingData, dec2018, by = "Code")
newHousingData <- newHousingData[c(3,8:10,19,28)]
names(newHousingData)[names(newHousingData)=="State.x"] <- "State"
names(newHousingData)[names(newHousingData)=="Region.x"] <- "Region"
names(newHousingData)[names(newHousingData)=="Difference.x"] <- "Decade99"
names(newHousingData)[names(newHousingData)=="Difference.y"] <- "Decade09"
names(newHousingData)[names(newHousingData)=="Difference"] <- "Decade18"
```
  
```{r}
newPopulation = populationChanges[c(2,11:13)]
question3Data = inner_join(newHousingData, newPopulation, by = "Code")
```
  
  
```{r fig.width=10 , fig.height=10, echo =FALSE}
library(gridExtra)
p1 = ggplot(question3Data, aes(x = changePercent1018, y = abs(Decade18))) + geom_point(alpha = 0.2) + geom_smooth(se = FALSE) + geom_text(aes(label = Code), size = 3) 
p2 = ggplot(question3Data, aes(x = changePercent0010, y = abs(Decade09))) + geom_point(alpha = 0.2) + geom_smooth() + geom_text(aes(label = Code))
p3 = ggplot(question3Data, aes(x = changePercent9000, y = abs(Decade99))) + geom_point(alpha = 0.2) + geom_smooth() + geom_text(aes(label = Code))

grid.arrange(p1, p2, p3, nrow = 3, ncol = 1)

```
  
```{r}
q3RegionalData = gather(question3Data, key = "Decade", value = "HousingPrices",
       Decade99, Decade09, Decade18)
q3RegionalData = gather(q3RegionalData, key = "Years", value = "PopulationChanges",
       changePercent9000, changePercent0010, changePercent1018)
```
  


```{r}
x = q3RegionalData[q3RegionalData$State == 'California', ]
x
```




```{r fig.height=12, fig.width=12}
ggplot(q3RegionalData, aes(x = PopulationChanges, y = HousingPrices, color = Decade)) + geom_point() + facet_wrap(State~.)
```

-

```{r fig.height=8, fig.width=12}
ggplot(question3Data, aes(x = PopulationChanges, y = HousingPrices, color = Decade)) + geom_point() + geom_smooth(se = FALSE, method = 'loess', degree = 0.001) + facet_grid(Region~.)
```











    ggtitle("Changes in Population by Decade") +
    xlab("State") +
    ylab("Population Change")
```


```{r}
ggplot(populationChanges, aes(x =  reorder(Code, changePercent9000), y = value, color = ColorsIndicate)) + 
    geom_point(aes(y = changePercent9000, col = "changePercent9000")) + 
    geom_point(aes(y = changePercent0010, col = "changePercent0010")) +
    geom_point(aes(y = changePercent1018, col = "changePercent1018")) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Percent Changes in Population by Decade") +
    xlab("State") +
    ylab("Population Change")
```

```{r, fig.height=8}
ggplot(populationChanges, aes(x = reorder(Region, change0010), y = value, color = ColorsIndicate)) +
    geom_point(aes(y = change9000, col = "change9000")) + 
    geom_point(aes(y = change0010, col = "change0010")) +
    geom_point(aes(y = change1018, col = "change1018")) +
  facet_wrap(Code ~ .) +
  #geom_smooth(method="gam", formula = y~s(x), se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Population Density Change by State") +
  xlab("State") +
  ylab("Population Change")
```

```{r, fig.height=8}
ggplot(populationChanges, aes(x = reorder(Region, change0010), y = value, color = ColorsIndicate)) +
    geom_point(aes(y = changePercent9000, col = "changePercent9000")) + 
    geom_point(aes(y = changePercent0010, col = "changePercent0010")) +
    geom_point(aes(y = changePercent1018, col = "changePercent1018")) +
  facet_wrap(Code ~ .) +
  #geom_smooth(method="gam", formula = y~s(x), se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Percent Population Density Change by State") +
  xlab("State") +
  ylab("Population Change")
```













```{r}
housingData = housePriceData[c(1:4,8,9)]

jan1990 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 1990, ]
dec1999 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 1999, ]
dec1999$HousingPriceChange1999 = dec1999$Price - jan1990$Price
dec1999 = dec1999[c(3,8:10)]

jan2000 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 2000, ]
dec2009 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 2009, ]
dec2009$HousingPriceChange2009 = dec2009$Price - jan2000$Price
dec2009 = dec2009[c(3,8:10)]

jan2010 = housePriceData[housePriceData$Month == 1 & housePriceData$Year == 2010, ]
dec2018 = housePriceData[housePriceData$Month == 12 & housePriceData$Year == 2018, ]
dec2018$HousingPriceChange2018 = dec2018$Price - jan2010$Price
dec2018 = dec2018[c(3,8:10)]

housingPriceChange = inner_join(dec1999, dec2009, by = "Code")
housingPriceChange = inner_join(housingPriceChange, dec2018, by = "Code")
housingPriceChange = housingPriceChange[c(1,4,7:10)]

populationChanges$PopulationChange1999 = populationChanges$Population2000 - populationChanges$Population1990
populationChanges$PopulationChange2009 = populationChanges$Population2010 - populationChanges$Population2000
populationChanges$PopulationChange2018 = populationChanges$Population2018 - populationChanges$Population2010
populationChanges = populationChanges[-c(4:7)]

populationHousingData = inner_join(housingPriceChange, populationChanges, by = "Code")
populationHousingData = populationHousingData[-c(7, 8)]
populationHousingData = populationHousingData[c(1, 4, 5, 2, 3, 6:9)]

populationHousingData1999 = populationHousingData[c(1:4, 7)]
populationHousingData1999$Year = 1999
names(populationHousingData1999)[names(populationHousingData1999)=="HousingPriceChange1999"] <- "HousingPriceChange"
names(populationHousingData1999)[names(populationHousingData1999)=="PopulationChange1999"] <- "PopulationChange"

populationHousingData2009 = populationHousingData[c(1:3, 5, 8)]
populationHousingData2009$Year = 2009
names(populationHousingData2009)[names(populationHousingData2009)=="HousingPriceChange2009"] <- "HousingPriceChange"
names(populationHousingData2009)[names(populationHousingData2009)=="PopulationChange2009"] <- "PopulationChange"

populationHousingData2018 = populationHousingData[c(1:3, 6, 9)]
populationHousingData2018$Year = 2018
names(populationHousingData2018)[names(populationHousingData2018)=="HousingPriceChange2018"] <- "HousingPriceChange"
names(populationHousingData2018)[names(populationHousingData2018)=="PopulationChange2018"] <- "PopulationChange"

populationHousingData = populationHousingData1999
populationHousingData = rbind(populationHousingData, populationHousingData2009)
populationHousingData = rbind(populationHousingData, populationHousingData2018)

library(viridis)
#cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

 
```

```{r}
ggplot(populationHousingData, aes(x = log(abs(PopulationChange)), y = HousingPriceChange, group = Year, color = Year)) +
  geom_point() +
  geom_smooth(method = "rlm", se = FALSE) + 
  ggtitle("Population Change Vs. Housing Price Change") +
  xlab("Population Change") +
  ylab("Housing Price Change") +
#  scale_color_manual(values = cb_palette) +
  scale_color_viridis() +
  facet_wrap(Region.x ~ .)
```