---
title: "2018 vote switchers: Getting started"
author: "S470/670"
date: "Last updated `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
set.seed(470670)
```

Load libraries:

```{r, message = FALSE}
library(tidyverse)
```

Load data:

```{r}
DFP = read.csv("DFP_WTHH_release.csv")
DFP = as_tibble(DFP)
```

```{r, eval = FALSE}
names(DFP)
```

```{r}
n = nrow(DFP)
n
```

## Reproducing toplines

We want to check that we can get topline statistics that are close to the percentages published by Data for Progress here:

https://wthh.dataforprogress.org/get-the-data

Try it for the Age variable, without survey weights:

```{r}
group_by(DFP, age5) %>% summarise(Proportion = n()/n)
```

These are way off. Instead, use weights:

```{r}
group_by(DFP, age5) %>% summarise(`Weighted proportion` = sum(weight_DFP)/n)
```

This is still a bit different from the numbers given by Data for Progress (because survey weights are more complex than we pretend they are in this class), but it's much closer.

Try a weighted summary for `house3`, the main variable for 2018 vote:

```{r}
group_by(DFP, house3) %>% summarise(`Weighted proportion` = sum(weight_DFP)/n)
```

Again, it's not exactly right, but it's close.



## Vote switchers

Vote switchers can be determined from the following variables:

presvote16post: 1: Clinton, 2: Trump, 3-6: Other, 7: DNV

house3: 1: Dem, 2: GOP, 3: other/DNV

Non-Clinton to Dem:

```{r}
NC2D = (DFP$presvote16post != 1 & DFP$house3 == 1)
table(NC2D, useNA = "ifany")
```

Non-Trump to Rep:

```{r}
NT2R = (DFP$presvote16post != 2 & DFP$house3 == 2)
table(NT2R, useNA = "ifany")
```

Vote switchers are non-Clinton to Dem or non-Trump to Rep:

```{r}
switcher = (NC2D | NT2R)
table(switcher, useNA = "ifany")
```

Subset the vote switchers:

```{r}
switchers = DFP[switcher != FALSE,]
nrow(switchers)
```

The 465 vote-switchers include 291 for whom NC2D was true, 171 for whom NT2R was true, and 3 NA to Dem or Rep.


